---
title: "Energy Mix Per Capita in Indonesia"
subtitle: "Revised - Problem Set 2 | SIS 750"
author: "Abhishu Karki"
date: "Apr 21, 2026"
format:
  html:
    self-contained: true
    theme: flatly
    toc: false
    fig-width: 11
    fig-height: 5.5
    page-layout: full
execute:
  echo: false
  warning: false
  message: false
---


```{r}
#| label: setup

library(tidyverse)
library(readxl)
library(plotly)
library(scales)
```

```{r}
#| label: data-prep

# Data----
Percapenergy <- read_excel("per_capita_energ_stacked.xlsx")

Indoenergy <- Percapenergy |>
  filter(Entity == 'Indonesia') |>
  mutate(
    Total_energy = (Coal + Oil + Gas + Hydropower + Wind + Solar + Other_renewables),
    .after = 9
  )

# Wrangle----
energy_source <- c("Coal", "Oil", "Gas", "Hydropower", "Wind", "Solar", "Other_renewables")

energy_df <- Indoenergy |>
  select("Year", all_of(energy_source)) |>
  filter(Year >= 1985 & Year <= 2024)

barplot_data <- energy_df |>
  pivot_longer(cols = -Year, names_to = "Source", values_to = "Share") |>
  mutate(
    Share = as.numeric(format(Share, scientific = FALSE)),
    Source = factor(Source, levels = energy_source)
  )

# Renewables----
renewables_data <- barplot_data |>
  filter(Source %in% c("Hydropower","Solar", "Wind", "Other_renewables"), Year >= 2018)
```

```{r}
#| label: colours

energy_colors <- c(
  Coal= "#2c3e50",Oil = "#a0392c",Gas = "#e67e22",Hydropower = "#2980b9", Wind = "#a48ce6",Solar = "#f1c40f",Other_renewables = "#27ae60"
  )

energy_labels <- c(
  Coal = "Coal",
  Oil = "Oil",
  Gas = "Natural Gas",
  Hydropower = "Hydropower",
  Wind = "Wind",
  Solar = "Solar",
  Other_renewables = "Other renewables"
)
```



```{r}
#| label: fig-a
#| fig-width: 6
#| fig-height: 6
#| fig-cap: "Full energy mix, 1985–2024"

# Panel A-----
stacked <- barplot_data |>
  ggplot(aes(
    x = Year,y = Share, fill = Source,
    text = paste0(
      "<b>", energy_labels[as.character(Source)], "</b><br>",
      "Year: ", Year, "<br>",
      "Energy: ", comma(round(Share, 1)), " kWh/person"
    )
  )) +
  geom_bar(width = 1, stat = "identity", position = "stack", alpha = 0.95) +
  scale_fill_manual(values = energy_colors, labels = energy_labels) +
  labs(
    x = "Year",
    fill = "Energy source",
    y = "Per capita primary energy (kWh)"
  ) +
  scale_x_continuous(
    limits = c(1984.5, 2024.5),
    breaks = seq(1985, 2026, 5),
    expand = expansion(mult = c(0, 0))
  ) +
  scale_y_continuous(
    limits = c(0, 11000),
    labels = comma,
    expand = expansion(mult = c(0, 0))
  ) +
  theme_light() +
  theme(
    panel.grid.minor = element_blank(),
    plot.background = element_rect(fill = "#f2f2f2", color = NA),
    panel.grid.major.x = element_blank(),
    axis.title = element_blank(),
    legend.position = "right"
  )

#Interactive

ggplotly(stacked, tooltip = "text") |>
  layout(
    title = list(
      text = paste0(
        "<b>Per capita primary energy consumption by source in Indonesia</b>",
        "<br>",
        "<span style='font-size:18px; color:grey'>",
        "Primary energy is measured in kilowatt-hours per person <i>(Y axis)</i> per year <i>(X axis)</i></span>"
      ),
      x = 0,
      xanchor = "left",
      xref = "paper",
      font = list(size = 24, family = "sans")
    ),
    annotations = list(
      list(
        text = "<i>Data source: Energy Institute – Statistical Review of World Energy (2025)</i>",
        x = 1, y = -0.12,
        xref = "paper", yref = "paper",
        xanchor = "right", yanchor = "top",
        showarrow = FALSE,
        font = list(size = 12, color = "grey40", family = "sans"))
    ),
    legend = list(
      orientation = "v",
      x = 1.008, y = 0.5,
      font = list(size = 11, family = "verdana"),
      title = list(text = "<b>Energy source</b>")
    ),
    margin = list(t = 100, b = 80, l = 60, r = 150),
    hoverlabel = list(
      bgcolor = "#1a1a2e",
      font    = list(color = "white", size = 12)
    )
  )


```

## About this chart

Indonesia's per capita energy consumption grew from roughly 2,500 kWh in 1984 to over 10,000 kWh in 2024, with coal responsible for the majority of that increase. Rather than diversifying its energy mix as consumption rose, Indonesia became more reliant on fossil fuels, we see this in @fig-a with oil plateauing around 2004, gas growing consistently, and coal continuing to climb as the primary energy source. @fig-b narrows the focus to 2018–2024 to assess renewable progress, showing that solar and wind together account for under 20 kWh per person which is less than 0.2% of total consumption. While the larger renewable block consisting mainly of [biomass,hass and geothermal ](https://ourworldindata.org/grapher/per-capita-energy-stacked?country=~IDN&overlay=sources) has been growing significantly, Indonesia's efforts are not enough. Against Indonesia's stated targets of [23% renewables by 2030—already delayed from 2025—and net zero by 2060,](https://ieefa.org/resources/after-years-energy-crossroads-can-indonesia-pivot-2026) the data suggest a significant implementation gap as overall energy demand is still rising. The renewable capacity must outpace a growing baseline, not just substitute an existing one for the policy goals to be achieved.

```{r}
#| label: fig-b
#| fig-width: 6
#| fig-height: 4
#| fig-cap: "Renewable growth, 2018–2024"

#Panel B-----
renewables_plot <- renewables_data |>
    ggplot(aes(
      x = Year,
      y = Share,
      color = Source,
      group = Source,
      text = paste0(
        "<b>", as.character(Source), "</b><br>",
        "Year: ", Year, "<br>",
        "Energy: ", comma(round(Share, 1)), " kWh/person"
      )
    )) +
    geom_line(linewidth = 0.5) +
    geom_point(size = 2.5) +
    scale_color_manual(
      values = c(Hydropower = "#2980b9", Solar = "#f1c40f", Wind = "#a48ce6", Other_renewables =
                   "#27ae60"),
      labels = energy_labels 
      )+
    scale_x_continuous(breaks = 2018:2024) +
    scale_y_continuous(labels = comma, expand = expansion(mult = c(0.02, 0.05))) +
    labs( color = "Source" ) +
    theme_light() +
    theme(
      plot.background = element_rect(fill = "#f2f2f2", color = NA),
      panel.grid.minor = element_blank(),
      panel.grid.major.x = element_blank(),
      axis.title = element_blank(),
      legend.position = "right"
    )

ggplotly(renewables_plot, tooltip = "text") |>
  layout(
    title = list(
      text = paste0(
        "<b>Renewable energy growth per capita in Indonesia</b><br>",
        "<span style='font-size:20px; color:grey'>",
        "Kilowatt-hours per person · 2018–2024</span>"
        ),
      x = 0,
      xanchor = "left",
      xref = "paper",
      font = list(size = 24, family = "sans")
      ),
    annotations = list(
      list(
        text = "<i>Data source: Energy Institute – Statistical Review of World Energy (2025)</i>",
        x = 1, y = -0.12,
        xref = "paper", yref = "paper",
        xanchor = "right", yanchor = "top",
        showarrow = FALSE,
        font = list(size = 12, color = "grey40", family = "sans")
      )),
    paper_bgcolor = "#f2f2f2", plot_bgcolor = "#f2f2f2",
    xaxis = list(
      showgrid = FALSE,
      title = ""
    ),
    yaxis = list(
      showgrid = FALSE,
      title = ""
    ),
    legend = list(
      orientation = "v",
      x = 1.008, y = 0.5,
      font = list(size = 11, family = "verdana"),
      title = list(text = "<b>Source</b>")
    ),
    margin = list(t = 100, b = 80, l = 60, r = 150),
    hoverlabel = list(
      bgcolor = "#1a1a2e",
      font    = list(color = "white", size = 12)
    ))
    
```

## Data Source
Energy Institute - Statistical Review of World Energy (2025); Population based on various sources (2024) – with major processing by Our World in Data. “Coal consumption per capita” [dataset]. Energy Institute, “Statistical Review of World Energy”; Various sources, “Population” [original data](https://ourworldindata.org/grapher/per-capita-energy-stacked?country=~IDN).